### 2020-11-11 tables.Rmd
not great

---
title: "Tables"
output: bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(kableExtra)
library(tidyverse)
library(ggplot2)
```

# Top visited private HS (2-mode network)

```{r}
df <- readRDS('./assets/tables/table_2mode.RDS')

knitr::kable(head(df, n = 20)) %>% kable_styling(latex_options = 'scale_down')
```

Table \@ref(tab:hs-print-privu) shows characteristics of the 20 private high schools that were visited by the most private colleges and universities

```{r, hs-print-privu, echo = FALSE}
df <- readRDS('./assets/tables/table_2mode.RDS')

knitr::kable(head(df, n = 20, caption = 'ego table 3')) %>% kable_styling(latex_options = 'scale_down')
```

# Aggregated visited private HS (2-mode network)

```{r}
df <- readRDS('./assets/tables/table_2mode_agg.RDS')

knitr::kable(df) %>% kable_styling(latex_options = 'scale_down')
```

# Ego network table

```{r}
df <- readRDS('./assets/tables/table_ego.RDS')
knitr::kable(df %>% select(-ID, -Characteristics), row.names = F) %>% kable_styling(latex_options = 'scale_down')
```

# Ego network barplot

```{r, fig.width=3.5, fig.height=3}
#ego_df <- readRDS('./assets/tables/table_ego_privu.RDS')
ego_df <- readRDS('./assets/tables/table_ego_pubu.RDS')
univ_order <- (ego_df %>% arrange(Cluster, Type, as.numeric(Rank), University))$University

plot_characteristic <- function(df, type, characteristics) {
  sub_df <- df %>% select(University, characteristics) %>%
    gather(key, value, characteristics) %>%
    mutate(value = parse_number(value))

  # colors <- sub_df$color[order(sub_df$University)]

  sub_df$University <- factor(sub_df$University, levels = rev(univ_order))

  ggplot(sub_df, aes(fill = key, y = value, x = University)) + 
    geom_bar(stat='identity', position = position_fill(reverse = TRUE)) +
    ggtitle('Title') +
    xlab('') + ylab('') +
    labs(fill = type) +
    scale_y_continuous(labels = scales::percent, expand = c(0, 0, 0.05, 0)) +
    theme(legend.position = 'top',
          legend.title = element_text(size = 6, face = 'bold'),
          legend.text = element_text(size = 6),
          legend.margin = margin(0, 0, 0, 0),
          legend.box.margin = margin(0, 0, -5, -30),
          legend.key.size = unit(0.3, 'cm'),
          panel.background = element_blank(),
          axis.ticks.y = element_blank(),
          # axis.text.y = element_text(color = rep(c('red', 'green', 'blue', 'yellow', 'black'), 5)),
          text = element_text(size = 8),
          plot.title = element_text(color = 'white')) +
    coord_flip()
}

pdf(str_c('assets/figures/', "ego_network_figure.pdf"), paper = "a4r")
par(mfrow=c(2, 2))
plot_characteristic(ego_df, 'Region', c('Northeast', 'Midwest', 'South', 'West'))
plot_characteristic(ego_df, 'Religion', c('Catholic', 'Conserv', 'Nonsect', 'Other'))
plot_characteristic(ego_df, '% Black/Latinx/Native', c('c1_lt10', 'c2_10to25', 'c3_25to50', 'c4_50+'))
plot_characteristic(ego_df, 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'))

dev.off()

```

```{r, fig.width=3.5, fig.height=3}
type_vec <- c("privu","pubu")
  #type_vec

for (i in 1:length(type_vec)) {

  #df_name <- str_c("table_ego_",type_vec[i],".RDS")
  ego_df <-readRDS(file.path('.','assets','tables',str_c("table_ego_",type_vec[i],".RDS")))
  #ego_df %>% glimpse() %>% print()

  univ_order <- (ego_df %>% arrange(Cluster, Type, as.numeric(Rank), University))$University

  pdf(file.path('.','assets','figures',str_c("ego_network_figure_",type_vec[i],".pdf")), paper = "a4r")

  plot_characteristic(ego_df, 'Region', c('Northeast', 'Midwest', 'South', 'West'))
  plot_characteristic(ego_df, 'Religion', c('Catholic', 'Conserv', 'Nonsect', 'Other'))
  plot_characteristic(ego_df, '% Black/Latinx/Native', c('c1_lt10', 'c2_10to25', 'c3_25to50', 'c4_50+'))
  plot_characteristic(ego_df, 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'))

  dev.off()
}

```

# Ego network graph

See Figure \@ref(fig:ego-plot).

```{r ego-plot, fig.align = 'center', out.width = '100%', fig.cap = 'Ego network'}
knitr::include_graphics('assets/figures/plot_2mode_privu.pdf')
```