### 2021-05-17 text/tables_figures_may_2021.Rmd
draft changes revising results

---
title: "Recruiting Chapter"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \floatsetup{capposition=top}
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)

library(kableExtra)
library(tidyverse)
library(gridExtra)
```

# Results

ADD TABLE OF UNIVERSITY CHARACTERISTICS HERE

## RQ1: Scale

```{r event-count-figure-privu, fig.align = 'center', echo = FALSE, fig.cap = "Number of events by type and in-state, out-of-state for public institutions"}
knitr::include_graphics('./../assets/figures/plot_event_count_pubu.pdf')
```

```{r event-count-figure-pubu, fig.align = 'center', echo = FALSE, fig.cap = "Number of events by type and in-state, out-of-state for private institutions"}
knitr::include_graphics('./../assets/figures/plot_event_count_privu.pdf')
```

## RQ2: Characteristics of high schools

```{r, include = FALSE}
# ggplot's default color palette
# library(scales)
# show_col(hue_pal()(4))

plot_characteristic <- function(df, type, characteristic, label_text, visit_count = FALSE) {

  if (visit_count) {
    df <- df %>% mutate(univ_name = str_c(univ_name, ' (N=', total, ')'))
  }

  # create variable for university region; and turn rank into a numeric variable
  df <- df %>% mutate(
    univ_region=factor(
      str_split(characteristics, pattern = '\\|', simplify=TRUE)[,1],levels = c('northeast','midwest','south','west')),
    rank = as.numeric(rank)
  )

  if (type == 'Region') {
    ordered_df <- df %>% arrange(univ_region, univ_name)
  } else {
    ordered_df <- df %>% arrange(rank, univ_name)
  }

  univ_order <- ordered_df$univ_name

  sub_df <- df %>% select(univ_name, all_of(characteristic)) %>%
    gather(key, value, characteristic) %>%
    mutate(value = parse_number(value))

  sub_df$univ_name <- factor(sub_df$univ_name, levels = rev(univ_order))
  sub_df$key <- factor(sub_df$key, levels = characteristic)

  ggplot(sub_df, aes(fill = key, y = value, x = univ_name)) + 
    geom_bar(stat='identity', position = position_fill(reverse = TRUE)) +
    ggtitle('Title') +
    xlab('') + ylab('') +
    labs(fill = type) +
    scale_fill_manual(labels = label_text, values = c('#F8766D', '#7CAE00', '#00BFC4', '#C77CFF')) +
    scale_y_continuous(labels = scales::percent, expand = c(0, 0, 0.05, 0)) +
    theme(legend.position = 'top',
          legend.title = element_text(size = 6, face = 'bold'),
          legend.text = element_text(size = 6),
          legend.margin = margin(0, 0, 0, 0),
          legend.box.margin = margin(0, 0, -5, -30),
          legend.key.size = unit(0.3, 'cm'),
          panel.background = element_blank(),
          axis.ticks.y = element_blank(),
          text = element_text(size = 8),
          plot.title = element_text(color = 'white')) +
    coord_flip()
}

```

```{r, fig.width=3.5, fig.height=3, echo = FALSE, message = FALSE, include = FALSE}
#all univ privu all_pubu_outst pub pubu_outst
# running one plot at a time

# region

  pdf(file.path('.','..', 'assets', 'figures', str_c('ego_network_figure_', 'region', '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 1))

  a <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'pubu', '.RDS'))),
    'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'), visit_count = TRUE
  )

  b <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'privu', '.RDS'))),
    'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'), visit_count = TRUE
  )

  grid.arrange(a, b, nrow=2, ncol=1)

  dev.off()

# religion

  pdf(file.path('.','..', 'assets', 'figures', str_c('ego_network_figure_', 'religion', '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 1))

  a <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'pubu', '.RDS'))),
    'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'), visit_count = TRUE
  )

  b <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'privu', '.RDS'))),
    'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'), visit_count = TRUE
  )

  grid.arrange(a, b, nrow=2, ncol=1)

  dev.off()

# racial composition

  pdf(file.path('.','..', 'assets', 'figures', str_c('ego_network_figure_', 'race', '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 1))

  a <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'pubu', '.RDS'))),
    '% Black/Latinx/Native', c('c1_lt10', 'c2_10to20', 'c3_20to50', 'c4_50+'), c('<10%', '10-20%', '20-50%', '50%+'), visit_count = TRUE
  )

  b <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'privu', '.RDS'))),
    '% Black/Latinx/Native', c('c1_lt10', 'c2_10to20', 'c3_20to50', 'c4_50+'), c('<10%', '10-20%', '20-50%', '50%+'), visit_count = TRUE
  )

  grid.arrange(a, b, nrow=2, ncol=1)

  dev.off()

# academic reputation

  pdf(file.path('.','..', 'assets', 'figures', str_c('ego_network_figure_', 'hs_rank', '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 1))

  a <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'pubu', '.RDS'))),
    'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'), visit_count = TRUE
  )

  b <- plot_characteristic(
    df = readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', 'privu', '.RDS'))),
    'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'), visit_count = TRUE
  )

  grid.arrange(a, b, nrow=2, ncol=1)

  dev.off()

```

\newpage

```{r ego-network-figure-region, eval = FALSE, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of visited private HS, region"}
knitr::include_graphics('./../assets/figures/ego_network_figure_region.pdf')
```

```{r, fig.width=3.5, fig.height=3, echo = FALSE, message = FALSE, eval = FALSE}

type_vec <- c('pubu', 'pubu_outst','privu','univ','univ_pubu_outst')
#type_vec <- c('privu', 'pubu', 'pubu_outst', 'all', 'all_pubu_outst')

for (i in 1:length(type_vec)) {
  ego_df <-readRDS(file.path('.','..', 'assets', 'tables', str_c('table_ego_', type_vec[i], '.RDS')))

  pdf(file.path('.','..', 'assets', 'figures', str_c('ego_network_figure_', type_vec[i], '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 2))

  a <- plot_characteristic(ego_df, 'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'), visit_count = TRUE)
  b <- plot_characteristic(ego_df, 'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'))
  c <- plot_characteristic(ego_df, '% Black/Latinx/Native', c('c1_lt10', 'c2_10to20', 'c3_20to50', 'c4_50+'), c('<10%', '10-20%', '20-50%', '50%+'), visit_count = TRUE)
  d <- plot_characteristic(ego_df, 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'))

  grid.arrange(a, b, c, d, nrow=2, ncol=2, widths=c(1, 0.92))
  dev.off()

  # Need to use print() for ggplot inside loop in order for it to appear
  # https://stackoverflow.com/a/15678596/6373540
  # print(a)
  # print(b)
  # print(c)
  # print(d)
}
```

## RQ3: Overlapping sets of high schools

# References

<div id="refs"></div>

<!-- TABLES -->

\newpage
<!-- private col/univ -->

```{r hs-print-privu, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_privu.RDS') %>% head(n=20)

knitr::kable(
  df %>% select(-name,-closeness), caption = 'Top 20 most central private HS visited by private institutions') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

```{r hs-table-privu-agg, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_agg_privu.RDS')

knitr::kable(df, caption = 'Characteristics of private HS visited by private institutions, by degree band') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

<!-- public research universities -->

```{r hs-print-pubu, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_pubu.RDS') %>% head(n=20)

knitr::kable(
  df %>% select(-name,-closeness), caption = 'Top 20 most central private HS visited by public institutions') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

```{r hs-table-pubu-agg, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_agg_pubu.RDS')

knitr::kable(df, caption = 'Characteristics of private HS visited by public institutions, by degree band') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

<!-- public and private col/univ -->

```{r hs-print-both, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_both.RDS') %>% head(n=20)

knitr::kable(
  df %>% select(-name,-closeness), caption = 'Top 20 most central private HS visited by all institutions') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

```{r hs-table-both-agg, echo = FALSE}
df <- readRDS('./../assets/tables/table_2mode_agg_both.RDS')

knitr::kable(df, caption = 'Characteristics of private HS visited by all institutions, by degree band') %>% 
  landscape(margin = NULL) %>%
  kable_styling(latex_options = 'scale_down')
```

<!-- FIGURES -->

```{r em-funnel, echo = FALSE, fig.align = 'center', fig.cap = "The enrollment funnel", out.width = "70%"}
knitr::include_graphics('./../assets/images/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

\newpage

```{r privhs-universe, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of universe of private HS and private HS that received at least 1 visit from sample institutions, unweighted and weighted by 12th grade enrollment"}
knitr::include_graphics('./../assets/figures/plot_pss_universe.pdf')
```

\begin{center}
* \textit{Universe of private HS (N=3673), private HS visited by sample institutions (N=1738); Universe of private HS is defined as any private schools that enrolls at least ten grade 12 students and is not a special education school, an alternative school, an early childhood center, or an independent school.}
\end{center}

\newpage

\newgeometry{top=1in, left=0cm, right=0cm, bottom=0cm}

\begin{landscape}

```{r two-mode-plot, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "2-mode network for all institutions, colored by geographic region"}
knitr::include_graphics('./../assets/figures/plot_2mode_all_region.pdf')

#https://stackoverflow.com/questions/15625990/how-to-set-size-for-local-image-using-knitr-for-markdown
```

```{r plot-1mode-privu, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "1-mode network for private institutions, colored by cluster"}
knitr::include_graphics('./../assets/figures/plot_1mode_privu.pdf')
```

\end{landscape}

\restoregeometry

\newgeometry{top=1in, left=0cm, right=0cm, bottom=0cm}

\newpage

```{r ego-network-figure-privu, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of private HS visited by private institutions"}
knitr::include_graphics('./../assets/figures/ego_network_figure_privu.pdf')
```

\begin{center}
* \textit{Universities are grouped by clusters, as indicated by alternating bolded and normal font}
\end{center}

\begin{landscape}

\newpage

```{r villanova-religion, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "Ego network of Villanova University, colored by religious affiliation"}
knitr::include_graphics('./../assets/figures/villanova_religion.pdf')
```

```{r emory-region, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "Ego network of Emory University, colored by geographic region"}
knitr::include_graphics('./../assets/figures/emory_region.pdf')
```

```{r plot-1mode-pubu, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "1-mode network for public institutions, colored by cluster"}
knitr::include_graphics('./../assets/figures/plot_1mode_pubu.pdf')
```

```{r plot-1mode-pubu-outst, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "1-mode network for public institutions and out-of-state visits only, colored by cluster", eval = FALSE}
knitr::include_graphics('./../assets/figures/plot_1mode_pubu_outst.pdf')
```

\end{landscape}

\restoregeometry

\newgeometry{top=1in, left=0cm, right=0cm, bottom=0cm}

\newpage

```{r ego-network-figure-pubu, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of private HS visited by public institutions"}
knitr::include_graphics('./../assets/figures/ego_network_figure_pubu.pdf')
```

\begin{center}
* \textit{Universities are grouped by clusters, as indicated by alternating bolded and normal font}
\end{center}

\begin{landscape}

```{r plot-1mode-all, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "1-mode network for all institutions, colored by cluster", include= FALSE, eval = FALSE}
knitr::include_graphics('./../assets/figures/plot_1mode_all.pdf')
```

```{r plot-1mode-all-pubu-outst, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "1-mode network for all institutions, out-of-state visits only for public universities, colored by cluster"}
knitr::include_graphics('./../assets/figures/plot_1mode_all_pubu_outst.pdf')
```

\end{landscape}

\restoregeometry

\newgeometry{top=1in, left=0cm, right=0cm, bottom=0cm}

\newpage

```{r ego-network-figure-all-pubu-outst, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of private HS visited by all institutions, out-of-state visits only for public universities"}
knitr::include_graphics('./../assets/figures/ego_network_figure_all_pubu_outst.pdf')
#out.width="1.1\\linewidth", 
```

\begin{center}
* \textit{Universities are grouped by clusters, as indicated by alternating bolded and normal font}
\end{center}