### 2021-06-11 text/tables_figures_june_2021.Rmd
add 1-mode tables

---
title: 'Recruiting Chapter'
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
  - \usepackage{pdflscape}
  - \usepackage{geometry}
  - \usepackage{floatrow}
  - \floatsetup{capposition=top}
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(scales)

knitr::opts_chunk$set(echo = F, message = F)

data_dir <- file.path('..', 'data')
tables_dir <- file.path('..', 'assets', 'tables')
figures_dir <- file.path('..', 'assets', 'figures')

univ_sample_df <- readRDS(file.path(data_dir, 'univ_sample.RDS')) %>%
  arrange(desc(classification), rank)
events_df <- readRDS(file.path(data_dir, 'events_data.RDS'))

pubhs_universe_df <- readRDS(file.path(data_dir, 'pubhs_universe.RDS'))
privhs_universe_df <- readRDS(file.path(data_dir, 'privhs_universe.RDS'))
```

# University characteristics

```{r}
univ_sample <- univ_sample_df %>% 
  select(classification, univ_abbrev, rank, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_p, pctfreshwh, pctfreshbl, pctfreshap, pctfreshhi, pctfreshal)

kable(univ_sample %>% arrange(desc(classification), rank)) %>%
  collapse_rows(columns = 1, latex_hline = 'major', valign = 'middle') %>%
  kable_styling(latex_options = 'scale_down')
```

# RQ1

```{r}
events_count <- events_df %>%
  mutate(event_type = case_when(
    event_type %in% c('pub_hs', 'priv_hs') ~ event_type,
    event_type == 'cc' & event_loc == 'instate' ~ event_type,
    TRUE ~ 'other'
  )) %>%
  select(univ_abbrev, event_loc, event_type) %>% 
  group_by(univ_abbrev, event_loc, event_type) %>% 
  summarise(count = n())

kable(events_count[order(match(events_count$univ_abbrev, univ_sample_df$univ_abbrev)), ], longtable = TRUE) %>%
  kable_styling(font_size = 7.1, latex_options = c('repeat_header'))
```

## Number of events by type, in-state vs. out-of-state

- `events_count_pubu.pdf`
- `events_count_privu.pdf`
- `events_count_privc.pdf`

```{r}
include_graphics(file.path(figures_dir, 'events_count_pubu.pdf'))
include_graphics(file.path(figures_dir, 'events_count_privu.pdf'))
include_graphics(file.path(figures_dir, 'events_count_privc.pdf'))
```

\newpage

## Number of events by public vs. private HS

- `events_hs_count_pubu.pdf`
- `events_hs_count_privu.pdf`
- `events_hs_count_privc.pdf`

```{r}
include_graphics(file.path(figures_dir, 'events_hs_count_pubu.pdf'))
include_graphics(file.path(figures_dir, 'events_hs_count_privu.pdf'))
include_graphics(file.path(figures_dir, 'events_hs_count_privc.pdf'))
```

\newpage

## Actual vs. proportional visited HS

- Below shows actual/proportional numbers for all vs. out-of-state HS visited
- Numbers differ from above plots because above shows event counts (some HS can be visited multiple times) whereas below numbers count each HS once (no matter how many times they're visited)
- % private is generally higher for out-of-state. If we include all HS (both in and out) the % actual private is usually lowered, as seen below

```{r}
for (i in 1:nrow(univ_sample_df)) {
  print(univ_sample_df[[i, 'univ_abbrev']])
  uid <- univ_sample_df[[i, 'univ_id']]
  ust <- univ_sample_df[[i, 'state_code']]

  visited_pub_hs <- (events_df %>% filter(univ_id == uid, event_type == 'pub_hs'))$school_id %>% unique()
  visited_pub_state <- (pubhs_universe_df %>% filter(ncessch %in% visited_pub_hs))$state_code %>% unique()
  universe_pub_hs <- (pubhs_universe_df %>% filter(state_code %in% visited_pub_state))$ncessch %>% unique()

  visited_pub_state_out <- (pubhs_universe_df %>% filter(ncessch %in% visited_pub_hs, state_code != ust))$state_code %>% unique()
  visited_pub_hs_out <- (pubhs_universe_df %>% filter(ncessch %in% visited_pub_hs, state_code != ust))$ncessch %>% unique()
  universe_pub_hs_out <- (pubhs_universe_df %>% filter(state_code %in% visited_pub_state_out))$ncessch %>% unique()

  visited_priv_hs <- (events_df %>% filter(univ_id == uid, event_type == 'priv_hs'))$school_id %>% unique()
  visited_priv_state <- (privhs_universe_df %>% filter(school_id %in% visited_priv_hs))$state_code %>% unique()
  universe_priv_hs <- (privhs_universe_df %>% filter(state_code %in% visited_priv_state))$school_id %>% unique()

  visited_priv_state_out <- (privhs_universe_df %>% filter(school_id %in% visited_priv_hs, state_code != ust))$state_code %>% unique()
  visited_priv_hs_out <- (privhs_universe_df %>% filter(school_id %in% visited_priv_hs, state_code != ust))$school_id %>% unique()
  universe_priv_hs_out <- (privhs_universe_df %>% filter(state_code %in% visited_priv_state_out))$school_id %>% unique()

  visited_hs <- c(visited_pub_hs, visited_priv_hs)
  visited_hs_out <- c(visited_pub_hs_out, visited_priv_hs_out)

  universe_hs <- c(universe_pub_hs, universe_priv_hs)
  universe_hs_out <- c(universe_pub_hs_out, universe_priv_hs_out)

  writeLines(str_c('Total HS visited (unique): ', length(visited_hs)))
  writeLines(str_c('  % private HS (actual): ', round(length(visited_priv_hs) / length(visited_hs) * 100, 1)))
  writeLines(str_c('  % private HS (proportional): ', round(length(universe_priv_hs) / length(universe_hs) * 100, 1)))

  writeLines(str_c('Total out-of-state HS visited (unique): ', length(visited_hs_out)))
  writeLines(str_c('  % private HS (actual): ', round(length(visited_priv_hs_out) / length(visited_hs_out) * 100, 1)))
  writeLines(str_c('  % private HS (proportional): ', round(length(universe_priv_hs_out) / length(universe_hs_out) * 100, 1)))
}
```

\newpage

# RQ2

## Characteristics of visited private HS

- `ego_network_region_pubu_privu_privc.pdf`
- `ego_network_religion_pubu_privu_privc.pdf`
- `ego_network_race_pubu_privu_privc.pdf`
- `ego_network_rank_pubu_privu_privc.pdf`
- `ego_network_enroll_pubu_privu_privc.pdf`

```{r}
include_graphics(file.path(figures_dir, 'ego_network_region_pubu_privu_privc.pdf'))
include_graphics(file.path(figures_dir, 'ego_network_religion_pubu_privu_privc.pdf'))
include_graphics(file.path(figures_dir, 'ego_network_race_pubu_privu_privc.pdf'))
include_graphics(file.path(figures_dir, 'ego_network_rank_pubu_privu_privc.pdf'))
include_graphics(file.path(figures_dir, 'ego_network_enroll_pubu_privu_privc.pdf'))
```

\newpage

# RQ3

## 1-mode plots

- `plot_1mode_pubu.pdf`
- `plot_1mode_privu.pdf`
- `plot_1mode_privc.pdf`
- `plot_1mode_u.pdf`, `plot_1mode_u_k5.pdf`
- `plot_1mode_all.pdf`, `plot_1mode_all_k5.pdf`, `plot_1mode_all_k6.pdf`

```{r}
include_graphics(file.path(figures_dir, 'plot_1mode_pubu.pdf'))
include_graphics(file.path(figures_dir, 'plot_1mode_privu.pdf'))
include_graphics(file.path(figures_dir, 'plot_1mode_privc.pdf'))

include_graphics(file.path(figures_dir, 'plot_1mode_u.pdf'))
include_graphics(file.path(figures_dir, 'plot_1mode_u_k5.pdf'))

include_graphics(file.path(figures_dir, 'plot_1mode_all.pdf'))
include_graphics(file.path(figures_dir, 'plot_1mode_all_k5.pdf'))
include_graphics(file.path(figures_dir, 'plot_1mode_all_k6.pdf'))
```

\newpage

## 1-mode tables

- `table_1mode_pubu.RDS`
- `table_1mode_privu.RDS`
- `table_1mode_privc.RDS`

Notes:

- Each .RDS stores a list containing a counts matrix (`onemode_df`) and percent matrix (`onemode_pct_df`)
- Some observations for the public university matrix and its corresponding 1-mode plot above
  - U of Alabama has a lot more visits to private HS than the rest and is in its own cluster as seen from plot
  - U of S.Carolina and CU Boulder are in a cluster together, and if we look at each of their columns in the pct matrix, they do have the most in common w/ each other (besides Alabama)
  - UNL, UC Riverside, and Stony Brook actually have very little commonly visited HS, but ended up being in the same cluster as seen from the plot. They do have the fewest number of private HS visits, so less likely to have visits in common w/ other univs. As seen from the plot, their vertices are located very far apart from each other and have no direct edges connecting them.

\newpage

\newgeometry{top=0.5cm, left=0in, right=0cm, bottom=0.5cm}

\begin{landscape}

```{r}
table_1mode_pubu <- readRDS(file = file.path(tables_dir, 'table_1mode_pubu.RDS'))

kable(table_1mode_pubu$onemode_df, caption = 'Public universities count matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(c(1, 15), width = '8.5em') %>% 
  column_spec(2:13, width = '7em') %>% 
  column_spec(c(14, 16), width = '6em')

kable(table_1mode_pubu$onemode_pct_df, caption = 'Public universities percent matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(c(1, 15), width = '8.5em') %>% 
  column_spec(2:13, width = '7em') %>% 
  column_spec(c(14, 16), width = '6em')

table_1mode_privu <- readRDS(file = file.path(tables_dir, 'table_1mode_privu.RDS'))

kable(table_1mode_privu$onemode_df, caption = 'Private universities count matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(1, width = '8.5em') %>%
  column_spec(2:15, width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em')

kable(table_1mode_privu$onemode_pct_df, caption = 'Private universities percent matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(1, width = '8.5em') %>%
  column_spec(2:15, width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em')

table_1mode_privc <- readRDS(file = file.path(tables_dir, 'table_1mode_privc.RDS'))

kable(table_1mode_privc$onemode_df, caption = 'Private colleges count matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(1, width = '8.5em') %>%
  column_spec(2:13, width = '7em') %>% 
  column_spec(11, width = '6em')

kable(table_1mode_privc$onemode_pct_df, caption = 'Private colleges percent matrix') %>%
  kable_styling(latex_options = 'scale_down') %>%
  column_spec(1, width = '8.5em') %>%
  column_spec(2:13, width = '7em') %>% 
  column_spec(11, width = '6em')
```

\end{landscape}