### 2023-11-09 soc_of_ed/soc_of_ed_proposal_jaquette.Rmd

---
#title: "The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities"
#author: 
#  - Ozan Jaquette
#  - Crystal Han
#  - Irma Castañeda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
  #bookdown::word_document2:
    toc: FALSE
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(scales)

knitr::opts_chunk$set(echo = F, message = F)

knitr::knit_hooks$set(inline = function(x) {   if(!is.numeric(x)){     x   }else{    prettyNum(round(x,2), big.mark=",")    } })

data_dir <- file.path('..', 'data')
tables_dir <- file.path('..', 'assets', 'tables')
figures_dir <- file.path('..', 'assets', 'figures')

univ_sample_df <- readRDS(file.path(data_dir, 'univ_sample.RDS')) %>%
  arrange(desc(classification), rank)
events_df <- readRDS(file.path(data_dir, 'events_data.RDS'))

pubhs_universe_df <- readRDS(file.path(data_dir, 'pubhs_universe.RDS'))
privhs_universe_df <- readRDS(file.path(data_dir, 'privhs_universe.RDS'))

ego_df <- readRDS(file.path(tables_dir, 'table_ego_all.RDS')) %>% 
  mutate(
    univ_region = str_split(characteristics, '\\|', simplify = T)[,1],
    univ_religion = str_split(characteristics, '\\|', simplify = T)[,2],
    univ_race = round(as.numeric(str_split(characteristics, '\\|', simplify = T)[,3]), 1),
    univ_rank = as.numeric(str_split(characteristics, '\\|', simplify = T)[,4]),
    univ_enroll = format(as.numeric(str_split(characteristics, '\\|', simplify = T)[,5]), big.mark = ',')
  )
```

**The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities**

\medskip

*During the 2017 calendar year, we tracked off-campus recruiting visits made by 14 selective private universities, 12 selective private liberal arts colleges, and 15 public research universities. We are still trying to figure out what happened.*
\medskip

**Introduction**. Harpeth Hall is a non-sectarian private school for girls located in the affluent Green Hills neighborhood of Nashville, Tennessee. Harpeth is an "A+" school ranked 337 in the nation according to Niche. Founded in 1951, the racial composition of Harpeth reflects broader regional patterns, where “efforts by the courts to desegregate public schools in the South resulted in white flight to private schools” [@RN4504, p. 14]. In Fall 2017, about 85% of Harpeth students were white and less than 5% were Black or Latinx.

From September 6th through October 17th, Harpeth received recruiting visits from 11 private universities in our sample (Northwestern, Boston College, Tufts, Villanova, Case Western Reserve, University of Denver, Emory, Tulane, Baylor, Texas Christian University), and three private liberal arts colleges (Harvey Mudd, Scrips, and Sewanee). Previous scholarship observes strong connections between private schools and selective private universities [@RN3519; @RN4407]. However, Harpeth also received visits from six out-of-state public research universities, the University of Cincinnati, University of South Carolina, the University of Alabama, University of Arkansas, University of Georgia, and CU Boulder.

Following the Supreme Court decision on race-conscious admissions, many universities have published statements about their commitment to diversity and racial justice. Talk is cheap. Drawing from theories of organizational behavior [@RN531; @RN959; @RN531], university recruiting behaviors that require significant expenditure are indicators of enrollment priorities. According to @RN4664, off-campus recruiting visits were the top-ranked recruiting expenditure item for both public and private universities. These visits leave a trail that problematizes the story told by university public relations efforts.

<!-- 
Therefore, analyses of which schools a university visits -- and which schools they ignore -- can yield insights about university enrollment goals.
What emerges is a story of white money and white privilege that flies in the face of  university public relations efforts.

Comparing the recruiting networks of public research universities and selective private universities yields novel insights about the competition for affluent, "full-pay" students.  
This research problematizes the story told by university public relations offices by analyzing off-campus recruiting visits to private high schools made by 14 selective private universities and 15 public research universities.

Tell an untold story that flies in the face of the story that university public relations efforts tells us
That the everyday recruiting practices of these universities actually serve the interests of white supremacy;
-->
This study analyzes off-campus recruiting visits to private high schools from a sample of 14 selective private universities and 15 public research universities. Recruiting visits indicate a social relationship, motivating social network analyses of network ties (visits) between actors (universities and schools): Analyses are informed by three research questions:
<!--

-->

1. How does the scale of visits to private schools by public research universities compare to the scale of visits to private schools by selective private universities?
1. How do the characteristics of private schools visited by public research universities compare to the characteristics of private schools visited by selective private universities?
1. To what extent do public and private univs visit overlapping sets of private schools?
<!--

-->

**Literature Review**. We bridge two disconnected literatures: (1) nonresident enrollment at public universities; and (2) linkages between private schools and private universities.
<!--

-->

@RN3753 found that public research universities responded to declines in state appropriations by growing enrollment from nonresident students, who pay higher prices than residents. @RN3685 shows that growth in the share of nonresident students is associated with declines in the share of Pell recipients and Black and LatinX students. Universities visit high schools to attract applicants and maintain relations with counselors. @RN4758 analyzed off-campus recruiting visits to public high schools by public research universities. Most universities made more visits to out-of-state high schools than in-state high schools. These out-of-state visits focused on affluent, predominantly white schools.

Scholarship from sociology has analyzed linkages between private schools and universities. Attending private school positively affects college access and college selectivity [e.g., @RN4642; @RN4614;@RN4625]. Although recently discovered by economics [@RN4935], sociology has long observed a special relationship between elite private schools and elite universities [@RN4407; @RN4647; @RN3519; @RN4640]. Moreover, sociology highlights the primacy of social relationships between elite schools and elite colleges.

Although these two literatures developed in isolation, they are parts of a broader story about school segregation and the competition for affluent students. In recent decades, Catholic school enrollment by middle-income families declined in the Midwest and the Northeast [@RN4504]. In the South, enrollment in Christian and nonsectarian private schools grew as a consequence of "white flight" from public schools. Historically, private schools were the domain of selective private universities. However, public research universities have an incentive to visit out-of-state private schools to find "full-pay" nonresident students. By analyzing the network of recruiting visits, we show whether public research universities and selective private universities are competing at similar sets of private schools.
<!-- 
 We also examine the characteristics of visited schools, which have implications for university enrollment composition.
-->

**Methods**. We collected off-campus recruiting visit data in 2017 by web-scraping events from university admissions websites. The analysis sample for this study consists of 15 public research universities (research-extensive) and 14 private universities ranked in the top 100 of US News National University rankings. We also incorporated secondary data from IPEDS, the American Community Survey, NCES Common Core of Data (public schools), and the Niche Best Private High Schools ranking. Off-campus recruiting visits are the network tie that defines a two-mode social network between universities (mode 1) and private schools (mode 2). We construct the two-mode network, transform the two-mode network to several one-mode networks, and utilize descriptive statistics to analyze these objects.

**Results**. Results for RQ1 (scale of visits to private schools) are shown in Figures \@ref(fig:events-hs-count-pubu), \@ref(fig:events-hs-count-privu), \@ref(fig:actual-proportional-pubu), and \@ref(fig:actual-proportional-privu). Results for RQ2 (characteristics of visited private schools) are shown in Figures \@ref(fig:region-pubu-privu), \@ref(fig:religion-pubu-privu), \@ref(fig:rank-pubu-privu), \@ref(fig:race-pubu-privu), and \@ref(fig:race-pubu-privhs-pubhs). Results for RQ3 (overlap in visited schools) are shown in Table \@ref(tab:privu-percent-visited-by-pubu-matrix) and in Figure \@ref(fig:plot-1mode-u), which shows the one-mode public university network with universities categorized into groups based on "community detection" cluster analysis methods.

**Significance**. Universities in our sample made a disproportionate number of visits to private schools. In turn, the disproportionate enrollment of private school students at these universities may be the consequence of intentional recruiting rather than  exogenous student demand. Visited private schools tend to be much whiter than visited public schools. Most public universities in our sample -- even those not in the South -- made a large share of visits in the South. In contrast to stated commitments about racial diversity, university recruiting behavior implicitly targets households that sought racially homogenous high schools.

\newpage

# References

<div id="refs"></div>

<!-- TABLES -->

\newpage

\clearpage

\begin{landscape}

```{r u-cluster1-percent-matrix, include=FALSE}
table_1mode_u <- readRDS(file = file.path(tables_dir, 'table_1mode_u.RDS'))

#kable(table_1mode_u$clusters_df[order(match(table_1mode_u$clusters_df$univ_abbrev, rownames(table_1mode_u$onemode_df))), ], caption = 'Clusters for public and private universities')

cluster_1 <- c(1, 4, 5, 7, 16, 18, 19, 24, 25, 28)
cluster_2 <- c(2, 3, 6, 8, 9, 10, 11, 12, 14, 15, 20, 22, 23)
cluster_3 <- c(13, 17, 21, 26, 27)
cluster_4 <- 29

kable(table_1mode_u$onemode_pct_df[cluster_1, cluster_1], caption = 'Public and private universities percent matrix (cluster 1)') %>%
  kable_styling(latex_options = 'scale_down')
```

```{r privu-percent-visited-by-pubu-matrix}
rownames(table_1mode_u$onemode_pct_df) <- colnames(table_1mode_u$onemode_pct_df)

kable(table_1mode_u$onemode_pct_df[c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29),c(2,3,4,5,6,8,10,11,12,14,15,19,20,26)], caption = 'For private universities, the percent of high schools visited by public universities') %>%
  kable_styling(latex_options = 'scale_down') %>%
  #column_spec(1, width = '8.5em') %>%
  column_spec(2:15, width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em')
```

\end{landscape}

\restoregeometry

<!-- FIGURES -->

\clearpage

\newpage

```{r events-hs-count-pubu, echo = FALSE, fig.align = 'left', out.width = "125%", fig.cap = "Number of visits to public and private high schools by public research universities"}
include_graphics(file.path(figures_dir, 'events_hs_count_pubu.pdf'))
```

\newpage

```{r events-hs-count-privu, echo = FALSE, fig.align = 'left', out.width = "125%", fig.cap = "Number of visits to public and private high schools by selective private universities"}
include_graphics(file.path(figures_dir, 'events_hs_count_privu.pdf'))
```

\newpage

```{r actual-proportional-pubu, echo = FALSE, fig.align = 'left', out.width = "125%", fig.cap = "Actual versus proportional number of private school visits, public research universities"}
include_graphics(file.path(figures_dir, 'events_hs_actual_proportional_pubu.pdf'))
```

\clearpage

```{r actual-proportional-privu, echo = FALSE, fig.align = 'left', out.width = "125%", fig.cap = "Actual versus proportional number of private school visits, selective private universities"}
include_graphics(file.path(figures_dir, 'events_hs_actual_proportional_privu.pdf'))
```

\clearpage

```{r region-pubu-privu, echo = FALSE, fig.cap = "Geographic region of visited private high schools"}
include_graphics(file.path(figures_dir, 'ego_network_region_pubu_privu.pdf'))
```

\newpage

```{r religion-pubu-privu, echo = FALSE, fig.cap = "Religious affiliation of visited private high schools"}
include_graphics(file.path(figures_dir, 'ego_network_religion_pubu_privu.pdf'))
```

\newpage

```{r rank-pubu-privu, echo = FALSE, fig.cap = "High school ranking of visited private high schools"}
include_graphics(file.path(figures_dir, 'ego_network_rank_pubu_privu.pdf'))
```

\newpage

```{r race-pubu-privu, echo = FALSE, fig.cap = "Percentage of students who identify as Black, Latinx, or Native at visited private high schools"}
include_graphics(file.path(figures_dir, 'ego_network_race_pubu_privu.pdf'))
```

\newpage

```{r race-pubu-privhs-pubhs, echo = FALSE, fig.cap = "Percentage of students who identify as Black, Latinx, or Native at visited private high schools vs. public high schools, public research universities"}
include_graphics(file.path(figures_dir, 'ego_network_race_pubu_privhs_pubhs.pdf'))
```

```{r plot-1mode-u, fig.align = 'center', echo = FALSE, out.width = "100%", fig.cap = "One-mode network for public and private universities, colored by cluster"}
knitr::include_graphics('./../assets/figures/plot_1mode_u.pdf')
```