### 2024-02-15 index.qmd
started revising soc of ed presentation

---
title: "The Private School Network"
subtitle: "Recruiting Visits to Private High Schools by Public and Private Universities"
bibliography: ./other/burd_recruiting_chapter.bib
format:
  revealjs: 
    theme: [default, ./assets/css/custom.scss]
    controls: true
    controls-layout: bottom-right
    controls-tutorial: true
    transition: slide
    background-transition: fade
    auto-stretch: false
    slide-level: 3
include-in-header: ./assets/html/header.html
# https://github.com/quarto-dev/quarto-cli/discussions/5853#discussioncomment-6126603
# filters:
#   - /path/to/relevel_headings.lua
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)  # For collapse_rows() to work: https://stackoverflow.com/a/67803915/6373540
library(readxl)

knitr::opts_chunk$set(echo = F, message = F)

data_dir <- file.path('.', 'data')
tables_dir <- file.path('.', 'assets', 'tables')
figures_dir <- file.path('.', 'assets', 'figures')

univ_sample_df <- readRDS(file.path(data_dir, 'univ_sample.RDS')) %>%
  arrange(desc(classification), rank)
events_df <- readRDS(file.path(data_dir, 'events_data.RDS'))

pubhs_universe_df <- readRDS(file.path(data_dir, 'pubhs_universe.RDS'))
privhs_universe_df <- readRDS(file.path(data_dir, 'privhs_universe.RDS'))

ego_df <- readRDS(file.path(tables_dir, 'table_ego_all.RDS')) %>% 
  mutate(
    univ_region = str_split(characteristics, '\\|', simplify = T)[,1],
    univ_religion = str_split(characteristics, '\\|', simplify = T)[,2],
    univ_race = round(as.numeric(str_split(characteristics, '\\|', simplify = T)[,3]), 1),
    univ_rank = as.numeric(str_split(characteristics, '\\|', simplify = T)[,4]),
    univ_enroll = format(as.numeric(str_split(characteristics, '\\|', simplify = T)[,5]), big.mark = ',')
  )
```

# Motivation

### Motivation

#### Off-campus recruiting visits to high schools in Nashville

CRYSTAL: INSERT MAP HERE

### Motivation

-   Recruiting visit indicates university and school have social relationship
-   Visits (edges) from university (mode 1) to high school (mode 2) yields a two-mode social network
-   Comparing recruiting network of public research and selective private universities gives insight about competition for affluent students

Research questions

1.  How does the scale of visits to private high schools vary for the public research universities and selective private universities?
2.  To what extent do public research universities and selective private universities visit overlapping sets of private high schools?
3.  What are the characteristics of private schools that receive visits from public research universities and selective private research universities?

# Literature Review

### Research on nonresident enrollment at public universities

- Increase nonresident enrollment in response to declines in state funding  [@RN3753]
- Nonresident students more affluent, less likely to be Black or Latinx [@RN3685]

Interventions to attract nonresident students

- Institutional financial aid [@RN3762]
- Recruiting to solicit 'inquiries' and applications
  - "Student list" purchases [@RN4985]
  - Off-campus recruiting visits [@RN4758; @RN4934]
    - Out-of state visits to public high schools focus on affluent public schools, avoid non-white communities

### Research on links between private schools and private universities

Studies find attending private school increases college access, selectivity

- Public school students: 5% attend "top 100" public univ; 2% attend top 100 private univ (Author calculations from HSLS 2009)
- Private school: 11% attend top 100 public; 17% attend top 100 private
- Nonsectarian private: 11% attend top 100 public; 40% attend top 100 private

Research from economics [@RN4935]

- Rich students have higher non-academic ratings because they attend elite private schools (extracurricular activities, recommendations)

Research from sociology highlights organizational linkages [@RN4935]

### slide title
#### slide sub-title

<hr>

Here's some text:

-   @RN4504 said blah...
-   Parenthetical citation [@RN4625].
-   **Bold** and *italic* text

More things:

-   Blah

# Methods

## Data and Sample

### Characteristics of universities in analysis sample

::: {#tbl-1 .table-wrapper}
```{r sample-characteristics}
univ_sample <- univ_sample_df %>% 
  filter(classification != 'private_libarts') %>% 
  select(classification, univ_abbrev, rank, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_p, pctfreshwh, pctfreshbl, pctfreshap, pctfreshhi, pctfreshal) %>% 
  mutate(
    classification = recode(
      classification,
      'public_research' = 'Public Research',
      'private_national' = 'Private National'
    ),
    satactcomp25 = format(round(satactcomp25), big.mark = ',', trim = T),
    satactcomp75 = format(round(satactcomp75), big.mark = ',', trim = T),
    tfuginst = str_c('$', format(round(tfuginst), big.mark = ',', trim = T)),
    tfugoutst = str_c('$', format(round(tfugoutst), big.mark = ',', trim = T)),
    ugftptfreshtot = format(round(ugftptfreshtot), big.mark = ',', trim = T),
    freshoutstpct = str_c(format(round(freshoutstpct * 100, 1), nsmall = 1), '%'),
    pgrnt_p = str_c(format(round(pgrnt_p, 1), nsmall = 1), '%'),
    pctfreshwh = str_c(format(round(pctfreshwh, 1), nsmall = 1), '%'),
    pctfreshbl = str_c(format(round(pctfreshbl, 1), nsmall = 1), '%'),
    pctfreshap = str_c(format(round(pctfreshap, 1), nsmall = 1), '%'),
    pctfreshhi = str_c(format(round(pctfreshhi, 1), nsmall = 1), '%'),
    pctfreshal = str_c(format(round(pctfreshal, 1), nsmall = 1), '%')
  ) %>% 
  arrange(desc(classification), rank)

names(univ_sample) <- c('Classification', 'University', 'Rank', '25th Percentile SAT/ACT Composite Score', '75th Percentile SAT/ACT Composite Score', 'In-State Tuition + Fees', 'Out-of-State Tuition + Fees', 'Total Enrolled Freshmen', 'Percent<br>Out-of-State Freshmen', 'Percent Pell Recipients', 'Percent White Freshmen', 'Percent Black Freshmen', 'Percent Asian Freshmen', 'Percent Hispanic Freshmen', 'Percent International Freshmen')

kbl(univ_sample, align = c('l', 'l', rep('r', 13)), escape = F) %>% 
  collapse_rows(columns = 1, valign = 'middle') %>% 
  column_spec(4, width_min = '1.2in') %>% 
  column_spec(6, width_min = '0.9in') %>% 
  column_spec(8, width_min = '0.9in')
```
:::

### Characteristics of universities in analysis sample compared to data collection sample

::: {.table-wrapper .freeze-col}
```{r sample-population-characteristics}
cc2000_edition_data <- read_excel(file.path(data_dir, 'cc2000_edition_data.xls'), sheet = 'Data')
pop_pubu <- (cc2000_edition_data %>% filter(`Control (text)` == 'Public', `CC2000 (text)` == 'Doctoral/Research Universities—Extensive'))$UNITID %>% as.character()

usnews_rankings <- read_csv(file.path(data_dir, 'usnews_rankings.csv'), col_select = c('source', 'control', 'rank', 'univ_id'))
pop_privu <- (usnews_rankings %>% filter(source == 'national-universities', control == 'private', rank <= 100))$univ_id %>% as.character()

sample_pubu <- (univ_sample_df %>% filter(classification == 'public_research'))$univ_id
sample_privu <- (univ_sample_df %>% filter(classification == 'private_national'))$univ_id

univ_pop <- read.csv(file.path(data_dir, 'ipeds_data.csv'), header = TRUE, na.strings = c('', 'NULL', 'NA'), stringsAsFactors = FALSE, colClasses = c('unitid' = 'character')) %>% as_tibble() %>%
  filter(endyear == 2017, unitid %in% c(pop_pubu, pop_privu, sample_pubu, sample_privu)) %>% 
  select(unitid, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_n, cohortsfaef, ugftptfreshwhmf, ugftptfreshblmf, ugftptfreshapmf, ugftptfreshhimf, ugftptfreshalmf, ugftptfreshmrmf, ugftptfreshnamf, ugftptfreshunmf, ugftptfreshasmf, ugftptfreshnhmf) %>% 
  mutate(
    pgrnt_p = pgrnt_n / cohortsfaef * 100,
    pctfreshwh = ugftptfreshwhmf / ugftptfreshtot * 100,
    pctfreshbl = ugftptfreshblmf / ugftptfreshtot * 100,
    pctfreshap = ugftptfreshapmf / ugftptfreshtot * 100,  # ap = as + nh
    pctfreshhi = ugftptfreshhimf / ugftptfreshtot * 100,
    pctfreshal = ugftptfreshalmf / ugftptfreshtot * 100,
    pctfreshmr = ugftptfreshmrmf / ugftptfreshtot * 100,
    pctfreshna = ugftptfreshnamf / ugftptfreshtot * 100,
    pctfreshun = ugftptfreshunmf / ugftptfreshtot * 100,
    pctfreshas = ugftptfreshasmf / ugftptfreshtot * 100,
    pctfreshnh = ugftptfreshnhmf / ugftptfreshtot * 100
  ) %>% 
  select(unitid, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_p, pctfreshwh, pctfreshbl, pctfreshap, pctfreshhi, pctfreshal)

format_data <- function(univ_ids) {
  apply(univ_pop %>% filter(unitid %in% univ_ids) %>% select(-unitid), 2, function(x) quantile(x, probs = c(0.25, 0.5, 0.75), na.rm = T)) %>% 
    as.data.frame() %>% 
    mutate(
      satactcomp25 = format(round(satactcomp25), big.mark = ',', trim = T),
      satactcomp75 = format(round(satactcomp75), big.mark = ',', trim = T),
      tfuginst = str_c('$', format(round(tfuginst), big.mark = ',', trim = T)),
      tfugoutst = str_c('$', format(round(tfugoutst), big.mark = ',', trim = T)),
      ugftptfreshtot = format(round(ugftptfreshtot), big.mark = ',', trim = T),
      freshoutstpct = str_c(format(round(freshoutstpct * 100, 1), nsmall = 1), '%'),
      pgrnt_p = str_c(format(round(pgrnt_p, 1), nsmall = 1), '%'),
      pctfreshwh = str_c(format(round(pctfreshwh, 1), nsmall = 1), '%'),
      pctfreshbl = str_c(format(round(pctfreshbl, 1), nsmall = 1), '%'),
      pctfreshap = str_c(format(round(pctfreshap, 1), nsmall = 1), '%'),
      pctfreshhi = str_c(format(round(pctfreshhi, 1), nsmall = 1), '%'),
      pctfreshal = str_c(format(round(pctfreshal, 1), nsmall = 1), '%')
    ) %>% 
    t()
}

univ_pop_df <- cbind(
  format_data(sample_pubu),
  format_data(pop_pubu),
  format_data(sample_privu),
  format_data(pop_privu)
)
rownames(univ_pop_df) <- c('25th Percentile SAT/ACT Composite Score', '75th Percentile SAT/ACT Composite Score', 'In-State Tuition + Fees', 'Out-of-State Tuition + Fees', 'Total Enrolled Freshmen', 'Percent Out-of-State Freshmen', 'Percent Pell Recipients', 'Percent White Freshmen', 'Percent Black Freshmen', 'Percent Asian Freshmen', 'Percent Hispanic Freshmen', 'Percent International Freshmen')
colnames(univ_pop_df) <- rep(c('25th percentile', '50th percentile', '75th percentile'), 4)

kbl(univ_pop_df, align = 'r', escape = F) %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=102)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=57)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>% 
  column_spec(1, bold = T) %>%
  column_spec(2:13, width_min = '0.9in')
```
:::

## Constructing Social Networks

## Analyses

# Results

## RQ1: Scale of Visits

### Number of HS visits by public research universities

![](./assets/figures/events_hs_count_pubu.png)

### Number of HS visits by selective private universities

![](./assets/figures/events_hs_count_privu.png)

### Proportional HS visits by public research universities

![](./assets/figures/events_hs_actual_proportional_pubu.png)

### Proportional HS visits by selective private universities

![](./assets/figures/events_hs_actual_proportional_privu.png)

## RQ2: Overlap of Visits

### One-mode percent matrix, all universities

CRYSTAL: CREATE ONE-MODE PERCENT MATRIX FOR ALL UNIVERSITIES; FREEZING PANES FOR FIRST COLUMN AND FIRST ROW

### One-mode network for all universities

![](./assets/figures/plot_1mode_u.png)

## RQ3: Characteristics of Visited Schools

### Order 2 ego network of University of Notre Dame

![](./assets/figures/nd_religion.png)

### Geographic region of visited private high schools

![](./assets/figures/ego_network_region_pubu_privu.png)

### Religious affiliation of visited private high schools

![](./assets/figures/ego_network_religion_pubu_privu.png)

### Racial composition of visited HS, public universities

![](./assets/figures/ego_network_race_pubu_privhs_pubhs.png)

### Racial composition of visited HS, private universities

![](./assets/figures/ego_network_race_privu_privhs_pubhs.png)

### High school ranking of visited private high schools

![](./assets/figures/ego_network_rank_pubu_privu.png)

# Discussion

### Slide w/ citation

#### Subtitle

<hr>

Here's some text:

-   @RN4504 said blah...
-   Parenthetical citation [@RN4625].
-   **Bold** and *italic* text

More things:

-   Blah

### References

::: {#refs}
:::

# Appendix

## RQ1: Scale of Visits

### All recruiting visits, public research universities

![](./assets/figures/events_count_pubu.png)

### All recruiting visits, public research universities

![](./assets/figures/events_count_privu.png)

## RQ2: Overlap of Visits

### One-mode public university percent matrix

::: {.table-wrapper .tbl-1mode}
```{r pubu-percent-matrix}
table_1mode_pubu <- readRDS(file = file.path(tables_dir, 'table_1mode_pubu.RDS'))
rownames(table_1mode_pubu$onemode_pct_df) <- colnames(table_1mode_pubu$onemode_pct_df)

names(table_1mode_pubu$onemode_pct_df) <- str_replace(names(table_1mode_pubu$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_pubu$onemode_pct_df, align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:16, width_min = '0.9in')
```
:::

### One-mode network, public universities

![](./assets/figures/plot_1mode_pubu.png)

### One-mode network, public universities

![](./assets/figures/plot_1mode_privu.png)

### One-mode private university percent matrix

::: {.table-wrapper .tbl-1mode}
```{r privu-percent-matrix}
table_1mode_privu <- readRDS(file = file.path(tables_dir, 'table_1mode_privu.RDS'))
rownames(table_1mode_privu$onemode_pct_df) <- colnames(table_1mode_privu$onemode_pct_df)

names(table_1mode_privu$onemode_pct_df) <- str_replace(names(table_1mode_privu$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_privu$onemode_pct_df, align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:15, width_min = '0.9in')
```
:::

### For private universities, the percent of high schools visited by public universities

::: {.table-wrapper .tbl-1mode}
```{r privu-percent-visited-by-pubu-matrix}
table_1mode_u <- readRDS(file = file.path(tables_dir, 'table_1mode_u.RDS'))
rownames(table_1mode_u$onemode_pct_df) <- colnames(table_1mode_u$onemode_pct_df)

names(table_1mode_u$onemode_pct_df) <- str_replace(names(table_1mode_u$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_u$onemode_pct_df[c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29),c(2,3,4,5,6,8,10,11,12,14,15,19,20,26)], align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:15, width_min = '0.9in')
```
:::

### For public universities, the percent of high schools visited by private universities

::: {.table-wrapper .tbl-1mode}
```{r pubu-percent-visited-by-privu-matrix}
kable(table_1mode_u$onemode_pct_df[c(2,3,4,5,6,8,10,11,12,14,15,19,20,26),c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29)], align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:16, width_min = '0.9in')
```
:::

## RQ3: Characteristics of Visited Schools

### Racial composition in the South, public universities

![](./assets/figures/south_race_pubu_privhs_pubhs.png)

### Racial composition in the South, private universities

![](./assets/figures/south_race_privu_privhs_pubhs.png)