### 2024-02-20 index.qmd
#13 add background options

---
title: "The Private School Network"
subtitle: "Recruiting Visits to Private High Schools by Public and Private Universities"
bibliography: ./other/burd_recruiting_chapter.bib
format:
  revealjs: 
    theme: [default, ./assets/css/custom.scss]
    controls: true
    controls-layout: bottom-right
    controls-tutorial: true
    transition: slide
    background-transition: fade
    auto-stretch: false
    slide-level: 3
    template-partials:
      - ./assets/html/title-slide.html
include-in-header: ./assets/html/header.html
# https://github.com/quarto-dev/quarto-cli/discussions/5853#discussioncomment-6126603
# filters:
#   - /path/to/relevel_headings.lua
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)  # For collapse_rows() to work: https://stackoverflow.com/a/67803915/6373540
library(readxl)

knitr::opts_chunk$set(echo = F, message = F)

data_dir <- file.path('.', 'data')
tables_dir <- file.path('.', 'assets', 'tables')
figures_dir <- file.path('.', 'assets', 'figures')

univ_sample_df <- readRDS(file.path(data_dir, 'univ_sample.RDS')) %>%
  arrange(desc(classification), rank)
events_df <- readRDS(file.path(data_dir, 'events_data.RDS'))

pubhs_universe_df <- readRDS(file.path(data_dir, 'pubhs_universe.RDS'))
privhs_universe_df <- readRDS(file.path(data_dir, 'privhs_universe.RDS'))

ego_df <- readRDS(file.path(tables_dir, 'table_ego_all.RDS')) %>% 
  mutate(
    univ_region = str_split(characteristics, '\\|', simplify = T)[,1],
    univ_religion = str_split(characteristics, '\\|', simplify = T)[,2],
    univ_race = round(as.numeric(str_split(characteristics, '\\|', simplify = T)[,3]), 1),
    univ_rank = as.numeric(str_split(characteristics, '\\|', simplify = T)[,4]),
    univ_enroll = format(as.numeric(str_split(characteristics, '\\|', simplify = T)[,5]), big.mark = ',')
  )
```

# Introduction

### Off-campus recruiting visits to high schools in Nashville {.caption}

<iframe src="./assets/maps/intro_map.html" width=100% height=100% allowtransparency="true"></iframe>

### Setup

Recruiting visit indicates university and school have social relationship

- The university wants to enroll some students from the school
- The school wants to send some students to the university

Visits from universities to schools yields a two-mode social network

- "Edges" = visits; "Vertices" = schools (mode 1) and universities (mode 2)

Research questions

1.  How does the scale of visits to private high schools vary for the public research universities and selective private universities?
2.  To what extent do public research universities and selective private universities visit overlapping sets of private high schools?
3.  What are the characteristics of private schools that receive visits from public research universities and selective private research universities?

# Literature Review

### Literature Review
#### Nonresident enrollment at public universities

- Increase nonresident enrollment in response to declines in state funding  [@RN3753]
- Nonresident students more affluent, less likely to be Black or Latinx [@RN3685]

Interventions to attract nonresident students

- Institutional financial aid to convert admits [@RN3762]
- Recruiting interventions at earlier stages of enrollment funnel
  - "Student list" purchases [@RN4985]
  - Off-campus recruiting visits [@RN4758; @RN4934]
    - Out-of state visits to public high schools focus on affluent public schools, avoid non-white communities

### Literature Review
#### Linkages between private schools and private universities

Studies find attending private school increases college access, selectivity

- Public school students: 5% attend "top 100" public univ; 2% attend top 100 private univ (Author calculations from HSLS 2009)
- Private school: 11% `@` top 100 public; 17% `@` top 100 private
- Nonsectarian private: 11% `@` top 100 public; 40% `@` top 100 private

@RN4935 study Ivy Plus admissions decisions

- Rich students have higher non-academic ratings because they attend elite private schools (extracurriculars, recommendations)

Sociology highlights school-college organizational linkages

- Visits maintain relations with counselors at feeder schools [@RN3519]
- Private school counselors lobby for marginal candidates [@RN4407]

### Literature Review
#### Synthesis

These two literatures have developed in isolation

- Pursuing out-of-state private school students by public research univs encroaches on historic territory of selective private univs

Analyzing network of recruiting visits to private schools gives insight about competition for affluent students

- Public and private universities may be targeting similar sets of schools
  - Direct competition or vying to be second choice for rejected applicant

Private school students more likely to be affluent, white than public school students [@RN4504]

- Competitive recruiting of private school students has downstream effects on racial and socioeconomic composition of colleges

# Methods

### Methods
#### Data and Sample

Data collection (see @RN4758)

- 2017 off-campus recruiting visits scraped from univ admissions websites

Data collection sample

- Public "research-extensive" universities from 2000 Carnegie (N=102)
- Private universities in top 100 of U.S. News *National Universities* (N=57)
- Private colleges in top 50 of *National Liberal Arts Colleges* (N=47)

Analysis sample: univs that posted complete recruiting visits on website

- 15 public research, 14 private universities, 12 private liberal arts (omitted)

Secondary data

- IPEDS; CCD; Private School Survey; Niche *Best Private High Schools*

### Characteristics of universities in analysis sample {.caption}

::: {#tbl-1 .table-wrapper}
```{r sample-characteristics}
univ_sample <- univ_sample_df %>% 
  filter(classification != 'private_libarts') %>% 
  select(classification, univ_abbrev, rank, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_p, pctfreshwh, pctfreshbl, pctfreshap, pctfreshhi, pctfreshal) %>% 
  mutate(
    classification = recode(
      classification,
      'public_research' = 'Public Research',
      'private_national' = 'Private National'
    ),
    satactcomp25 = format(round(satactcomp25), big.mark = ',', trim = T),
    satactcomp75 = format(round(satactcomp75), big.mark = ',', trim = T),
    tfuginst = str_c('$', format(round(tfuginst), big.mark = ',', trim = T)),
    tfugoutst = str_c('$', format(round(tfugoutst), big.mark = ',', trim = T)),
    ugftptfreshtot = format(round(ugftptfreshtot), big.mark = ',', trim = T),
    freshoutstpct = str_c(format(round(freshoutstpct * 100, 1), nsmall = 1), '%'),
    pgrnt_p = str_c(format(round(pgrnt_p, 1), nsmall = 1), '%'),
    pctfreshwh = str_c(format(round(pctfreshwh, 1), nsmall = 1), '%'),
    pctfreshbl = str_c(format(round(pctfreshbl, 1), nsmall = 1), '%'),
    pctfreshap = str_c(format(round(pctfreshap, 1), nsmall = 1), '%'),
    pctfreshhi = str_c(format(round(pctfreshhi, 1), nsmall = 1), '%'),
    pctfreshal = str_c(format(round(pctfreshal, 1), nsmall = 1), '%')
  ) %>% 
  arrange(desc(classification), rank)

names(univ_sample) <- c('Classification', 'University', 'Rank', '25th Percentile SAT/ACT Composite Score', '75th Percentile SAT/ACT Composite Score', 'In-State Tuition + Fees', 'Out-of-State Tuition + Fees', 'Total Enrolled Freshmen', 'Percent<br>Out-of-State Freshmen', 'Percent Pell Recipients', 'Percent White Freshmen', 'Percent Black Freshmen', 'Percent Asian Freshmen', 'Percent Hispanic Freshmen', 'Percent International Freshmen')

kbl(univ_sample, align = c('l', 'l', rep('r', 13)), escape = F) %>% 
  collapse_rows(columns = 1, valign = 'middle') %>% 
  column_spec(4, width_min = '1.4in') %>% 
  column_spec(6, width_min = '1in') %>% 
  column_spec(8, width_min = '1in')
```
:::

### Characteristics of universities in analysis sample compared to data collection sample {.caption}

::: {#tbl-2 .table-wrapper}
```{r sample-population-characteristics}
cc2000_edition_data <- read_excel(file.path(data_dir, 'cc2000_edition_data.xls'), sheet = 'Data')
pop_pubu <- (cc2000_edition_data %>% filter(`Control (text)` == 'Public', `CC2000 (text)` == 'Doctoral/Research Universities—Extensive'))$UNITID %>% as.character()

usnews_rankings <- read_csv(file.path(data_dir, 'usnews_rankings.csv'), col_select = c('source', 'control', 'rank', 'univ_id'))
pop_privu <- (usnews_rankings %>% filter(source == 'national-universities', control == 'private', rank <= 100))$univ_id %>% as.character()

sample_pubu <- (univ_sample_df %>% filter(classification == 'public_research'))$univ_id
sample_privu <- (univ_sample_df %>% filter(classification == 'private_national'))$univ_id

univ_pop <- read.csv(file.path(data_dir, 'ipeds_data.csv'), header = TRUE, na.strings = c('', 'NULL', 'NA'), stringsAsFactors = FALSE, colClasses = c('unitid' = 'character')) %>% as_tibble() %>%
  filter(endyear == 2017, unitid %in% c(pop_pubu, pop_privu, sample_pubu, sample_privu)) %>% 
  select(unitid, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_n, cohortsfaef, ugftptfreshwhmf, ugftptfreshblmf, ugftptfreshapmf, ugftptfreshhimf, ugftptfreshalmf, ugftptfreshmrmf, ugftptfreshnamf, ugftptfreshunmf, ugftptfreshasmf, ugftptfreshnhmf) %>% 
  mutate(
    pgrnt_p = pgrnt_n / cohortsfaef * 100,
    pctfreshwh = ugftptfreshwhmf / ugftptfreshtot * 100,
    pctfreshbl = ugftptfreshblmf / ugftptfreshtot * 100,
    pctfreshap = ugftptfreshapmf / ugftptfreshtot * 100,  # ap = as + nh
    pctfreshhi = ugftptfreshhimf / ugftptfreshtot * 100,
    pctfreshal = ugftptfreshalmf / ugftptfreshtot * 100,
    pctfreshmr = ugftptfreshmrmf / ugftptfreshtot * 100,
    pctfreshna = ugftptfreshnamf / ugftptfreshtot * 100,
    pctfreshun = ugftptfreshunmf / ugftptfreshtot * 100,
    pctfreshas = ugftptfreshasmf / ugftptfreshtot * 100,
    pctfreshnh = ugftptfreshnhmf / ugftptfreshtot * 100
  ) %>% 
  select(unitid, satactcomp25, satactcomp75, tfuginst, tfugoutst, ugftptfreshtot, freshoutstpct, pgrnt_p, pctfreshwh, pctfreshbl, pctfreshap, pctfreshhi, pctfreshal)

format_data <- function(univ_ids) {
  apply(univ_pop %>% filter(unitid %in% univ_ids) %>% select(-unitid), 2, function(x) quantile(x, probs = c(0.25, 0.5, 0.75), na.rm = T)) %>% 
    as.data.frame() %>% 
    mutate(
      satactcomp25 = format(round(satactcomp25), big.mark = ',', trim = T),
      satactcomp75 = format(round(satactcomp75), big.mark = ',', trim = T),
      tfuginst = str_c('$', format(round(tfuginst), big.mark = ',', trim = T)),
      tfugoutst = str_c('$', format(round(tfugoutst), big.mark = ',', trim = T)),
      ugftptfreshtot = format(round(ugftptfreshtot), big.mark = ',', trim = T),
      freshoutstpct = str_c(format(round(freshoutstpct * 100, 1), nsmall = 1), '%'),
      pgrnt_p = str_c(format(round(pgrnt_p, 1), nsmall = 1), '%'),
      pctfreshwh = str_c(format(round(pctfreshwh, 1), nsmall = 1), '%'),
      pctfreshbl = str_c(format(round(pctfreshbl, 1), nsmall = 1), '%'),
      pctfreshap = str_c(format(round(pctfreshap, 1), nsmall = 1), '%'),
      pctfreshhi = str_c(format(round(pctfreshhi, 1), nsmall = 1), '%'),
      pctfreshal = str_c(format(round(pctfreshal, 1), nsmall = 1), '%')
    ) %>% 
    t()
}

univ_pop_df <- cbind(
  format_data(sample_pubu),
  format_data(pop_pubu),
  format_data(sample_privu),
  format_data(pop_privu)
)
rownames(univ_pop_df) <- c('25th Percentile SAT/ACT Composite Score', '75th Percentile SAT/ACT Composite Score', 'In-State Tuition + Fees', 'Out-of-State Tuition + Fees', 'Total Enrolled Freshmen', 'Percent Out-of-State Freshmen', 'Percent Pell Recipients', 'Percent White Freshmen', 'Percent Black Freshmen', 'Percent Asian Freshmen', 'Percent Hispanic Freshmen', 'Percent International Freshmen')
colnames(univ_pop_df) <- rep(c('25th percentile', '50th percentile', '75th percentile'), 4)

kbl(univ_pop_df, align = 'r', escape = F) %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=102)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=57)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>% 
  column_spec(1, bold = T) %>%
  column_spec(2:13, width_min = '0.9in')
```
:::

### Methods
#### Constructing social networks

text
- Transform two-mode network to one-mode university network showing number of schools visited by both university $i$ and university $j$

### Methods
#### Analyses by research question

1.  How does the scale of visits to private high schools vary for the public research universities and selective private universities?
    - Simple descriptive statistics

\medskip

2.  To what extent do public research universities and selective private universities visit overlapping sets of private high schools?
    - Graph one-mode university network; universities allocated to categories based on cluster analysis

\medskip

3.  What are the characteristics of private schools that receive visits from public research universities and selective private research universities?
    - Descriptive statistics of characteristics of university order-1 "ego network" (schools the university visited):
      - Geographic region, religion, racial composition, academic reputation

# Results

## RQ1: Scale of Visits

### Number of HS visits by public research universities {.caption}

![](./assets/figures/events_hs_count_pubu.png)

### Number of HS visits by selective private universities {.caption}

![](./assets/figures/events_hs_count_privu.png)

### Proportional HS visits by public research universities {.caption}

![](./assets/figures/events_hs_actual_proportional_pubu.png)

### Proportional HS visits by selective private universities {.caption}

![](./assets/figures/events_hs_actual_proportional_privu.png)

## RQ2: Overlap of Visits

### One-mode percent matrix, all universities {.caption}

::: {.table-wrapper .tbl-1mode}
```{r pubu-privu-1mode}
table_1mode_u <- readRDS(file = file.path(tables_dir, 'table_1mode_u.RDS'))
rownames(table_1mode_u$onemode_pct_df) <- colnames(table_1mode_u$onemode_pct_df)

names(table_1mode_u$onemode_pct_df) <- str_replace(names(table_1mode_u$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_u$onemode_pct_df, align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:30, width_min = '0.9in')
```
:::

### One-mode network for all universities {.caption}

![](./assets/figures/plot_1mode_u.png)

## RQ3: Characteristics of Visited Schools

### Order 2 ego network of University of Notre Dame {.caption}

![](./assets/figures/nd_religion.png)

### Geographic region of visited private high schools {.caption}

![](./assets/figures/ego_network_region_pubu_privu.png)

### Religious affiliation of visited private high schools {.caption}

![](./assets/figures/ego_network_religion_pubu_privu.png)

### Racial composition of visited HS, public universities {.caption}

![](./assets/figures/ego_network_race_pubu_privhs_pubhs.png)

### Racial composition of visited HS, private universities {.caption}

![](./assets/figures/ego_network_race_privu_privhs_pubhs.png)

### High school ranking of visited private high schools {.caption}

![](./assets/figures/ego_network_rank_pubu_privu.png)

# Discussion

### Slide w/ citation

#### Subtitle

Here's some text:

-   @RN4504 said blah...
-   Parenthetical citation [@RN4625].
-   **Bold** and *italic* text

More things:

-   Blah

### References

::: {#refs}
:::

# Appendix

## RQ1: Scale of Visits

### All recruiting visits, public research universities {.caption}

![](./assets/figures/events_count_pubu.png)

### All recruiting visits, public research universities {.caption}

![](./assets/figures/events_count_privu.png)

## RQ2: Overlap of Visits

### One-mode public university percent matrix {.caption}

::: {.table-wrapper .tbl-1mode}
```{r pubu-percent-matrix}
table_1mode_pubu <- readRDS(file = file.path(tables_dir, 'table_1mode_pubu.RDS'))
rownames(table_1mode_pubu$onemode_pct_df) <- colnames(table_1mode_pubu$onemode_pct_df)

names(table_1mode_pubu$onemode_pct_df) <- str_replace(names(table_1mode_pubu$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_pubu$onemode_pct_df, align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:16, width_min = '0.9in')
```
:::

### One-mode network, public universities {.caption}

![](./assets/figures/plot_1mode_pubu.png)

### One-mode network, public universities {.caption}

![](./assets/figures/plot_1mode_privu.png)

### One-mode private university percent matrix {.caption}

::: {.table-wrapper .tbl-1mode}
```{r privu-percent-matrix}
table_1mode_privu <- readRDS(file = file.path(tables_dir, 'table_1mode_privu.RDS'))
rownames(table_1mode_privu$onemode_pct_df) <- colnames(table_1mode_privu$onemode_pct_df)

names(table_1mode_privu$onemode_pct_df) <- str_replace(names(table_1mode_privu$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_privu$onemode_pct_df, align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:15, width_min = '0.9in')
```
:::

### For private universities, the percent of high schools visited by public universities {.caption}

::: {.table-wrapper .tbl-1mode}
```{r privu-percent-visited-by-pubu-matrix}
table_1mode_u <- readRDS(file = file.path(tables_dir, 'table_1mode_u.RDS'))
rownames(table_1mode_u$onemode_pct_df) <- colnames(table_1mode_u$onemode_pct_df)

names(table_1mode_u$onemode_pct_df) <- str_replace(names(table_1mode_u$onemode_pct_df), ' \\(', '<br>(')

kable(table_1mode_u$onemode_pct_df[c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29),c(2,3,4,5,6,8,10,11,12,14,15,19,20,26)], align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:15, width_min = '0.9in')
```
:::

### For public universities, the percent of high schools visited by private universities {.caption}

::: {.table-wrapper .tbl-1mode}
```{r pubu-percent-visited-by-privu-matrix}
kable(table_1mode_u$onemode_pct_df[c(2,3,4,5,6,8,10,11,12,14,15,19,20,26),c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29)], align = 'r', escape = F) %>%
  column_spec(1, bold = T) %>% 
  column_spec(2:16, width_min = '0.9in')
```
:::

## RQ3: Characteristics of Visited Schools

### Racial composition in the South, public universities {.caption}

![](./assets/figures/south_race_pubu_privhs_pubhs.png)

### Racial composition in the South, private universities {.caption}

![](./assets/figures/south_race_privu_privhs_pubhs.png)